Eliminating prior-bias from sparse-projection tomographic reconstructions

### 2020-06-05 paper/tci.tex
retracted statement on CS; replaced "cs" by "l1-ls" in most places.

\bibliographystyle{IEEEtran}
\usepackage{lineno,hyperref}
\usepackage{amsmath,graphicx, amssymb,tabularx,booktabs}
\usepackage{subcaption}
\usepackage[font=small,skip=2pt]{caption}
\usepackage{boxedminipage}
\usepackage{algpseudocode}
\usepackage{algorithm}
\usepackage{xcolor}

@@

In this context, measurements in conventional data acquisition techniques are acquired by sampling the physical object uniformly above the Nyquist rate. However, in the last decade, reconstruction from sub-Nyquist sampled data has been made possible due to methods such as the widely used Compressed Sensing (CS) technique~\cite{Donoho,introCS}, which assumes the data to exhibit properties such as sparsity of the underlying image under certain mathematical transforms such as the Discrete Cosine Transform (DCT) or wavelet transforms. Fig.~\ref{fig:diff_methods} shows that although CS removes the artefacts created due to sub-sampling, its reconstruction is blurred. When prior information about the data is used along with CS, the reconstruction improves significantly. The algorithm presented in~\cite{Essam2015} drew inspiration from CS and solved for the reconstruction from limited views using alternating direction method of multipliers. In other techniques like in~\cite{Polak2017}, reconstruction is performed by estimating sparse gradients in the data. In~\cite{Li2015}, the required number of views is reduced by minimizing the total variation in the data. The measurements and an initial estimate of the data guide the iterative routines.

\begin{figure}[!h]
\centering
\fcolorbox{red}{white}{\begin{subfigure}[b]{0.155\linewidth}
        \includegraphics[width=1\textwidth]{../images/humerus/humerus_template1.png}
        \caption{}
     \end{subfigure}
\begin{subfigure}[b]{0.155\linewidth}
        \includegraphics[width=1\textwidth]{../images/humerus/humerus_template2.png}
        \caption{}
     \end{subfigure}
\begin{subfigure}[b]{0.155\linewidth}
        \includegraphics[width=1\textwidth]{../images/humerus/humerus_template3.png}
        \caption{}
     \end{subfigure}
\begin{subfigure}[b]{0.155\linewidth}
        \includegraphics[width=1\textwidth]{../images/humerus/humerus_template4.png}
        \caption{}
     \end{subfigure}
\begin{subfigure}[b]{0.155\linewidth}
        \includegraphics[width=1\textwidth]{../images/humerus/humerus_template5.png}
        \caption{}
     \end{subfigure}
\begin{subfigure}[b]{0.155\linewidth}
        \includegraphics[width=1\textwidth]{../images/humerus/humerus_template6.png}
        \caption{}
     \end{subfigure}}\\
\begin{subfigure}[b]{0.235\linewidth}
        \fcolorbox{yellow}{yellow}{\includegraphics[width=\textwidth]{../images/humerus/global_local/groundTruth.png}}
        \caption{Test}
\end{subfigure}
\quad
    \begin{subfigure}[b]{0.23\linewidth}
        \includegraphics[width=\textwidth]{../images/humerus/fbp_humerus.png}
        \caption{FBP}
     \end{subfigure}
    \begin{subfigure}[b]{0.23\linewidth}
        \includegraphics[width=\textwidth]{../images/humerus/global_local/reconNoPrior.png}
        \caption{CS}
    \end{subfigure}
    \begin{subfigure}[b]{0.23\linewidth}
        \includegraphics[width=\textwidth]{../images/humerus/global_local/global.png}
        \caption{This paper}
     \end{subfigure}
    \caption{\textit{Knowing more about the data helps.} Reconstructions of (g) a test slice  of size $(200,200)$ from the Humerus CT dataset~\cite{humerus} are performed using object-prior [(a)-(f)] and from measurements obtained from only 10 views. (h) Filtered backprojection (FBP) shows streaky sub-sampling artefacts (i) CS reduces the artefacts significantly while slightly blurring details (j) Prior information coupled with CS greatly improves the reconstruction.}
\label{fig:diff_methods}
\end{figure}

In addition, when some extra information about the object's structure is known~\cite{PICCS,cardiacPICCS,lubner2011,pirple,mota2017}, it is beneficial in further reducing the number of measurements.  These prior-based techniques use information from previously scanned data (`object-prior') of the same object or a similar one, and utilize it to reconstruct new volumes from very few sets of additional measurements. However, the above methods have the following two limitations: (1) the prior information may potentially overwhelm the essential details that appear in the test, (2) there is the key issue of choice of \textit{a particular} previously scanned object among the many previously acquired scans. A large part of this paper is dedicated to alleviating the former limitation, which is an important issue that has so far been overlooked in the literature on tomography. The latter limitation was relaxed in~\cite{liu2016,Xu2012} by building dictionary-based priors from multiple previously scanned objects. However, as reported in~\cite{my_dicta_paper}, dictionary priors are not as accurate and fast as global eigenspace priors.  Global eigenspace priors are better able to exploit the similarity of a test volume to object-prior, by assuming that the new test volume lies within the space spanned by the eigenvectors of the multiple representative previously scanned objects. Fig.~\ref{fig:diff_methods} shows the advantage of combining the global prior with CS.
\subsection{\textbf{Contributions}}
\label{sec:contributions}
In this work (Fig.~\ref{fig:prior_overview}), we focus on further reducing the number of measurements, with  particular emphasis on longitudinal studies. For example, we consider medical datasets which consist of multiple CT scans being taken during a radio-frequency ablation procedure~\cite{Dong2015}. The process consists of inserting a needle probe into a patient's body. When the needle reaches the tumor site, an electric current is passed to burn the tumor. Repeated CT scans of the patient need to be acquired in order to visualize the movement of the needle and to ensure that it is reaching the appropriate target. A few initial densely-sampled scans are used as object-prior to help the physician know the position of the needle. The later scans are used to reveal the exact changes during and after burning of the tumor (ablation).  In this context, we demonstrate the combined use of CS with the global prior, in two flavors: the vanilla (unselective) and selective global prior-based reconstruction. The choice of the number of measurement views and the type of reconstruction -- unselective or selective, is driven by the goal of the procedure.

@@

%--------------------RFA 2 few views
\begin{figure}[!h]
\centering
\subcaptionbox{Test}{\fcolorbox{white}{green}{\includegraphics[width=0.23\columnwidth]{../images/tmh/RFA2/very_few_views/colorTestIm.png}}}\hfill
\subcaptionbox{Backprojection}{\includegraphics[width=0.24\columnwidth]{../images/tmh/RFA2/very_few_views/fbp.png}}\hfill
\subcaptionbox{CS}{\includegraphics[width=0.24\columnwidth]{../images/tmh/RFA2/very_few_views/cs.png}}\hfill
\subcaptionbox{Proposed\\ method: unselective prior}{\includegraphics[width=0.24\columnwidth]{../images/tmh/RFA2/very_few_views/plain_pca.png}}
\caption[Representative results-1]{\textbf{Goal: Track new changes.}\small{ Reconstruction of slice 7 (`test') of Fig.~\ref{fig:RFA2_test_object-prior} from only 90 views, using (b) Filtered backprojection and no prior resulting in streaks, SSIM $= 0.48$ (c) CS and no prior resulting in blurred bone structures, SSIM $=0.35$ and (d) unselective global prior (slices 1-6 of Fig.~\ref{fig:RFA2_test_object-prior} are used as object-prior) resulting in clear bone structures with less streaks, SSIM $=0.55$. The region enclosed in red rectangle is our RoI as it contains both the new position of the needle and some background. All SSIM values are computed for this RoI.}}
\label{fig:RFA2_very_few_views}
\end{figure}

@@

\begin{figure}[!h]
\centering
\subcaptionbox{Test}{\fcolorbox{white}{blue}{\includegraphics[width=0.23\columnwidth]{../images/tmh/RFA2/few_views/colorTestIm.png}}}\hfill
\subcaptionbox{Backprojection}{\includegraphics[width=0.24\columnwidth]{../images/tmh/RFA2/few_views/fbp.png}}\hfill
\subcaptionbox{CS}{\includegraphics[width=0.24\columnwidth]{../images/tmh/RFA2/few_views/cs_dct.png}}\hfill
\subcaptionbox{Proposed\\ method: unselective\\ prior}{\includegraphics[width=0.24\columnwidth]{../images/tmh/RFA2/few_views/plain_pca.png}}
\subcaptionbox{Proposed\\ method: selective prior}{\includegraphics[width=0.24\columnwidth]{../images/tmh/RFA2/few_views/weighted_pca_all_methods_kk_0_01.png}}
\subcaptionbox{Weights map}{\includegraphics[width=0.275\columnwidth]{../images/tmh/RFA2/few_views/weights_colormap.png}}
\caption[Representative results-2]{\textbf{Goal: Observe details of new changes.} Reconstruction of slice 8 of Fig.~\ref{fig:RFA2_test_object-prior} from 120 views, using (b) FBP with SSIM $= 0.50$ (c) CS with SSIM $=0.46$ and (d) unselective global prior, with SSIM $=0.51$ (\emph{notice dominance of the prior: a prominent residual shadow of the needle which was present in the object-prior, but not present in the test image}), and (e) selective prior with SSIM $=0.56$ (\emph{notice that the dominance of the prior is significantly controlled}). The region enclosed in green rectangle is our Region of Interest (RoI) as it contains both the new position of the needle and some background. The SSIM is computed in this RoI. (f) shows the computed weights map (defined later in the paper) used for reconstruction. Darker intensities indicate lower weights to prior as these are the regions of new changes.}
\label{fig:RFA2_few_views}
\end{figure}

@@

To begin with, when an object is scanned multiple times, a set of high quality reconstructions (i.e. reconstructions from a dense set of projection views) may be chosen as object-prior for the reconstruction of future scan volumes, which in turn, may be scanned using far fewer measurements. The eigenspace $E_{\text{high}}$ of the $L$ previously scanned objects $Q_1,Q_2,...,Q_L$ is pre-computed. Here, it is assumed that the test volume can be expressed as a sparse linear combination of the principal components (eigenvectors of the covariance matrix) obtained from a group of structurally similar volumes. Hence, the object-prior is represented by means of PCA. For the eigenspace to encompass a range of possible structures in the test slice, the object-prior must represent a wide structural range. Moreover, if these volumes are not aligned, then they must be first registered before computing the prior. The prior is built by computing the covariance matrix from the template set $\{Q_i\}_{i=1}^L$. \emph{The space spanned by the eigenvectors $\{\boldsymbol{V_k}\}_{k=1}^{L-1}$ (eigenspace) of the covariance matrix is the global prior and is assumed to contain any test slice that is similar, but not necessarily identical to the object-prior.}  We use all of the $L-1$ orthogonal eigenvectors as a basis to represent the unknown test volume. Let $\boldsymbol{x} = \boldsymbol{\Upsilon\theta}$ denote the reconstructed volume, $\boldsymbol{y} = \boldsymbol{\mathcal{R} x}$ its measured tomographic projections, $\boldsymbol{\mathcal{R}}$ the tomographic projection operator, $\boldsymbol{\theta}$ the sparse coefficients of $\boldsymbol{x}$, $\boldsymbol{\Upsilon}$ the basis in which $\boldsymbol{x}$ is assumed to be sparse, $\boldsymbol{\mu}$ the mean of the previously scanned objects, and $\boldsymbol{\alpha}$ the vector of eigen-coefficients of the test scan, of which $\alpha_k$ is the $k^{\textrm{th}}$ element. Then, once the eigenspace is pre-computed, the test is reconstructed by minimizing the following cost function:
\vspace{4mm}
 \begin{equation}
   \setlength{\belowdisplayskip}{0pt} \setlength{\belowdisplayshortskip}{0pt}
\setlength{\abovedisplayskip}{-2pt} \setlength{\abovedisplayshortskip}{-2pt}
J_1(\boldsymbol{\theta,\alpha}) = \lVert\boldsymbol{\mathcal{R} x}- \boldsymbol{y}\rVert_2^2  + \lambda_1\lVert\boldsymbol{\theta}\rVert_1+\lambda_2\lVert\boldsymbol{x} - (\boldsymbol{\mu} + \sum_{k}\boldsymbol{V_k}\alpha_k)\rVert_2^2.
\label{Eq:main_prior}
\end{equation}
Here, $\lambda_1$, $\lambda_2$ are tunable weights given to the sparsity and prior terms respectively. The unknowns $\boldsymbol{\theta}$ and $\boldsymbol{\alpha}$
are solved by alternately minimizing $J_{\boldsymbol{\alpha}}(\boldsymbol{\theta})$ using a fixed $\boldsymbol{\alpha}$, and $J_{\boldsymbol\theta}(\boldsymbol{\alpha})$ using the resultant $\boldsymbol{\theta}$, where 
\begin{align}
J_{\boldsymbol{\alpha}}(\boldsymbol{\theta}) &\triangleq \lVert\boldsymbol{\mathcal{R} x- y}\rVert_2^2  + \lambda_1\lVert\boldsymbol{\theta}\rVert_1+\lambda_2\lVert\boldsymbol{x} - (\boldsymbol{\mu + V\alpha})\rVert_2^2, \\
J_{\boldsymbol\theta}(\boldsymbol{\alpha}) &\triangleq \lVert\boldsymbol{\Upsilon\theta} - (\boldsymbol{\mu + V\alpha})\rVert_2^2.
\end{align}
Note that $\boldsymbol{\theta}$ is solved for using the basis pursuit CS solver~\cite{l1ls}. Solving for $\boldsymbol{\alpha}$ leads to the closed form update:
\begin{equation}
\boldsymbol{\boldsymbol{\alpha}} = \boldsymbol{V}^T(\boldsymbol{\Upsilon \theta} -\boldsymbol{\mu}).
\end{equation}
 Optimal values of $\lambda_1, \lambda_2$ must be empirically chosen \textit{a priori}, based on the reconstructions of one of the template volumes (see also Sec.~\ref{sec:discussion}). The cost function described in Eq.~\ref{Eq:main_prior} is biconvex and the convergence of this optimization is guaranteed by the monotone convergence theorem.
 %---------------------------------------------------------
 \vspace{2mm}

@@

%-----------------------------potato 2D weights
\begin{figure}[!h]
    \begin{subfigure}[b]{0.24\linewidth}
        \includegraphics[width=\textwidth]{../images/potato/2D/weightsIm_fbp30.png}
        \caption{FBP}
    \end{subfigure}
    \begin{subfigure}[b]{0.24\linewidth}
        \includegraphics[width=\textwidth]{../images/potato/2D/weightsIm_cs_dct30.png}
        \caption{CS-DCT}
     \end{subfigure}
    \begin{subfigure}[b]{0.24\linewidth}
        \includegraphics[width=\textwidth]{../images/potato/2D/weightsIm_cs_wavelet30.png}
        \caption{CS-Haar}
     \end{subfigure}
    \begin{subfigure}[b]{0.24\linewidth}
        \includegraphics[width=\textwidth]{../images/potato/2D/weightsIm_art30.png}
        \caption{ART}
     \end{subfigure}
    \begin{subfigure}[b]{0.24\linewidth}
        \includegraphics[width=\textwidth]{../images/potato/2D/weightsIm_sart30.png}
        \caption{SART}
     \end{subfigure}
    \begin{subfigure}[b]{0.24\linewidth}
        \includegraphics[width=\textwidth]{../images/potato/2D/weightsIm_sirt30.png}
        \caption{SIRT}
     \end{subfigure}
    \begin{subfigure}[b]{0.24\linewidth}
        \includegraphics[width=\textwidth]{../images/potato/2D/weightsIm_all_methods30.png}
        \caption{combined}
     \end{subfigure}
      \caption{Weights-maps (corresponding to the difference between pilot reconstruction of the image in the last sub-figure of Fig.~\ref{fig:object-prior_test_potato_A1} and its projection onto the eigenspace $V_{low}$) constructed using different reconstruction methods, as specified in Eq.~\ref{eq:weightsEq}. The weights-maps are different because the reconstruction artefacts of the new structures in test image will be different for every reconstruction method used, as seen in Fig.~\ref{fig:reconstructions_diff_methods}. }
\label{fig:weights_map_2Dpotato}
\end{figure}
%-----------------------------potato 2D weights
\begin{figure}[!h]
    \begin{subfigure}[b]{0.24\linewidth}
        \includegraphics[width=\textwidth]{../images/potato/2D/fbp.png}
        \caption{FBP}
    \end{subfigure}
    \begin{subfigure}[b]{0.24\linewidth}
        \includegraphics[width=\textwidth]{../images/potato/2D/cs_dct.png}
        \caption{CS-DCT}
     \end{subfigure}
    \begin{subfigure}[b]{0.24\linewidth}
        \includegraphics[width=\textwidth]{../images/potato/2D/cs_wavelet.png}
        \caption{CS-Haar}
     \end{subfigure}
    \begin{subfigure}[b]{0.24\linewidth}
        \includegraphics[width=\textwidth]{../images/potato/2D/art.png}
        \caption{ART}
     \end{subfigure}
    \begin{subfigure}[b]{0.24\linewidth}
        \includegraphics[width=\textwidth]{../images/potato/2D/sart.png}
        \caption{SART}
     \end{subfigure}
    \begin{subfigure}[b]{0.24\linewidth}
        \includegraphics[width=\textwidth]{../images/potato/2D/sirt.png}
        \caption{SIRT}
     \end{subfigure}
    \begin{subfigure}[b]{0.24\linewidth}
        \includegraphics[width=\textwidth]{../images/potato/2D/weighted_pca_all_methods30.png}
        \caption{combined}
     \end{subfigure}
      \caption{(a)-(f):Different reconstructions of~\ref{fig:potato_test}. The magnitude and sharpness of the artefacts is different for each method. (g) Selective-prior method that combines weights-map information from all other methods. The SSIM of these reconstructions is shown in Table.~\ref{table:potato_2D_ssim}.} 
\label{fig:reconstructions_diff_methods}
\end{figure}

>>>>

\bibliographystyle{IEEEtran}
\usepackage{lineno,hyperref}
\usepackage{amsmath,graphicx, amssymb,tabularx,booktabs}
\usepackage{subcaption}
\usepackage[font=small,skip=2pt]{caption}
\usepackage{boxedminipage}
%\usepackage{algpseudocode}
%\usepackage{algorithm}
\usepackage{xcolor}

@@

In this context, measurements in conventional data acquisition techniques are acquired by sampling the physical object uniformly above the Nyquist rate. However, in the last decade, reconstruction from sub-Nyquist sampled data has been made possible due to methods such as the widely used Compressed Sensing (CS) technique~\cite{Donoho,introCS}, which assumes the data to exhibit properties such as sparsity of the underlying image under certain mathematical transforms such as the Discrete Cosine Transform (DCT) or wavelet transforms. Among the many optimizers used in CS, we used the l1-regularized least squares (l1-ls) form and observed that although this optimizer removes most of the artefacts created due to sub-sampling, its reconstruction is blurred when the number of views is very limited, as shown in Fig.~\ref{fig:diff_methods}. Specifically, the reconstruction in Fig.~\ref{fig:humerus_cs} is obtained by minimizing the following cost using the solver in~\cite{l1ls}:
 \begin{equation}
J_{CS}(\boldsymbol{\theta}) = \lVert\boldsymbol{\mathcal{R} x}- \boldsymbol{y}\rVert_2^2  + \lambda_1\lVert\boldsymbol{\theta}\rVert_1
\label{Eq:simple_cs}
\end{equation}
Here, $\boldsymbol{x} = \boldsymbol{\Upsilon\theta}$ denotes the reconstructed volume, $\boldsymbol{y}$ its measured tomographic projections, $\boldsymbol{\mathcal{R}}$ the tomographic projection operator such that ideally $\boldsymbol{y} = \boldsymbol{\mathcal{R} x}$, and $\boldsymbol{\theta}$ the sparse coefficients of $\boldsymbol{x}$. $\boldsymbol{\Upsilon}$ is the basis in which $\boldsymbol{x}$ is assumed to be sparse, and is chosen to be DCT here.
In addition to this sparsity prior, when object specific prior information is  used as well, the reconstruction improves significantly. The algorithm presented in~\cite{Essam2015} drew inspiration from CS and solved for the reconstruction from limited views using alternating direction method of multipliers. In other techniques like in~\cite{Polak2017}, reconstruction is performed by estimating sparse gradients in the data. In~\cite{Li2015}, the required number of views is reduced by minimizing the total variation in the data. The measurements and an initial estimate of the data guide the iterative routines.

\begin{figure}[!h]
\centering
\fcolorbox{red}{white}{\begin{subfigure}[b]{0.155\linewidth}
        \includegraphics[width=1\textwidth]{../images/humerus/humerus_template1.png}
        \caption{}
     \end{subfigure}
\begin{subfigure}[b]{0.155\linewidth}
        \includegraphics[width=1\textwidth]{../images/humerus/humerus_template2.png}
        \caption{}
     \end{subfigure}
\begin{subfigure}[b]{0.155\linewidth}
        \includegraphics[width=1\textwidth]{../images/humerus/humerus_template3.png}
        \caption{}
     \end{subfigure}
\begin{subfigure}[b]{0.155\linewidth}
        \includegraphics[width=1\textwidth]{../images/humerus/humerus_template4.png}
        \caption{}
     \end{subfigure}
\begin{subfigure}[b]{0.155\linewidth}
        \includegraphics[width=1\textwidth]{../images/humerus/humerus_template5.png}
        \caption{}
     \end{subfigure}
\begin{subfigure}[b]{0.155\linewidth}
        \includegraphics[width=1\textwidth]{../images/humerus/humerus_template6.png}
        \caption{}
     \end{subfigure}}\\
\begin{subfigure}[b]{0.235\linewidth}
        \fcolorbox{yellow}{yellow}{\includegraphics[width=\textwidth]{../images/humerus/global_local/groundTruth.png}}
        \caption{Test}
\end{subfigure}
\quad
    \begin{subfigure}[b]{0.23\linewidth}
        \includegraphics[width=\textwidth]{../images/humerus/fbp_humerus.png}
        \caption{FBP}
     \end{subfigure}
    \begin{subfigure}[b]{0.23\linewidth}
        \includegraphics[width=\textwidth]{../images/humerus/global_local/reconNoPrior.png}
        \caption{l1-ls}
        \label{fig:humerus_cs}
    \end{subfigure}
    \begin{subfigure}[b]{0.23\linewidth}
        \includegraphics[width=\textwidth]{../images/humerus/global_local/global.png}
        \caption{This paper}
     \end{subfigure}
    \caption{\textit{Knowing more about the data helps.} Reconstructions of (g) a test slice  of size $(200,200)$ from the Humerus CT dataset~\cite{humerus} are performed using object-prior [(a)-(f)] and from measurements obtained from only 10 views. (h) Filtered backprojection (FBP) shows streaky sub-sampling artefacts (i) l1-ls reduces the artefacts significantly while slightly blurring details (j) Prior information coupled with l1-ls (CS) greatly improves the reconstruction.}
\label{fig:diff_methods}
\end{figure}

In addition, when some extra information about the object's structure is known~\cite{PICCS,cardiacPICCS,lubner2011,pirple,mota2017}, it is beneficial in further reducing the number of measurements.  These prior-based techniques use information from previously scanned data (`object-prior') of the same object or a similar one, and utilize it to reconstruct new volumes from very few sets of additional measurements. However, the above methods have the following two limitations: (1) the prior information may potentially overwhelm the essential details that appear in the test, (2) there is the key issue of choice of \textit{a particular} previously scanned object among the many previously acquired scans. A large part of this paper is dedicated to alleviating the former limitation, which is an important issue that has so far been overlooked in the literature on tomography. The latter limitation was relaxed in~\cite{liu2016,Xu2012} by building dictionary-based priors from multiple previously scanned objects. However, as reported in~\cite{my_dicta_paper}, dictionary priors are not as accurate and fast as global eigenspace priors.  Global eigenspace priors are better able to exploit the similarity of a test volume to object-prior, by assuming that the new test volume lies within the space spanned by the eigenvectors of the multiple representative previously scanned objects. Fig.~\ref{fig:diff_methods} shows the advantage of combining the global prior with the l1-ls form of CS.
\subsection{\textbf{Contributions}}
\label{sec:contributions}
In this work (Fig.~\ref{fig:prior_overview}), we focus on further reducing the number of measurements, with  particular emphasis on longitudinal studies. For example, we consider medical datasets which consist of multiple CT scans being taken during a radio-frequency ablation procedure~\cite{Dong2015}. The process consists of inserting a needle probe into a patient's body. When the needle reaches the tumor site, an electric current is passed to burn the tumor. Repeated CT scans of the patient need to be acquired in order to visualize the movement of the needle and to ensure that it is reaching the appropriate target. A few initial densely-sampled scans are used as object-prior to help the physician know the position of the needle. The later scans are used to reveal the exact changes during and after burning of the tumor (ablation).  In this context, we demonstrate the combined use of l1-ls (CS) with the global prior, in two flavors: the vanilla (unselective) and selective global prior-based reconstruction. The choice of the number of measurement views and the type of reconstruction -- unselective or selective, is driven by the goal of the procedure.

@@

%--------------------RFA 2 few views
\begin{figure}[!h]
\centering
\subcaptionbox{Test}{\fcolorbox{white}{green}{\includegraphics[width=0.23\columnwidth]{../images/tmh/RFA2/very_few_views/colorTestIm.png}}}\hfill
\subcaptionbox{Backprojection}{\includegraphics[width=0.24\columnwidth]{../images/tmh/RFA2/very_few_views/fbp.png}}\hfill
\subcaptionbox{l1-ls}{\includegraphics[width=0.24\columnwidth]{../images/tmh/RFA2/very_few_views/cs.png}}\hfill
\subcaptionbox{Proposed\\ method: unselective prior}{\includegraphics[width=0.24\columnwidth]{../images/tmh/RFA2/very_few_views/plain_pca.png}}
\caption[Representative results-1]{\textbf{Goal: Track new changes.}\small{ Reconstruction of slice 7 (`test') of Fig.~\ref{fig:RFA2_test_object-prior} from only 90 views, using (b) Filtered backprojection and no prior resulting in streaks, SSIM $= 0.48$ (c) l1-ls resulting in blurred bone structures, SSIM $=0.35$ and (d) unselective global prior (slices 1-6 of Fig.~\ref{fig:RFA2_test_object-prior} are used as object-prior) resulting in clear bone structures with less streaks, SSIM $=0.55$. The region enclosed in red rectangle is our RoI as it contains both the new position of the needle and some background. All SSIM values are computed for this RoI.}}
\label{fig:RFA2_very_few_views}
\end{figure}

@@

\begin{figure}[!h]
\centering
\subcaptionbox{Test}{\fcolorbox{white}{blue}{\includegraphics[width=0.23\columnwidth]{../images/tmh/RFA2/few_views/colorTestIm.png}}}\hfill
\subcaptionbox{Backprojection}{\includegraphics[width=0.24\columnwidth]{../images/tmh/RFA2/few_views/fbp.png}}\hfill
\subcaptionbox{l1-ls}{\includegraphics[width=0.24\columnwidth]{../images/tmh/RFA2/few_views/cs_dct.png}}\hfill
\subcaptionbox{Proposed\\ method: unselective\\ prior}{\includegraphics[width=0.24\columnwidth]{../images/tmh/RFA2/few_views/plain_pca.png}}
\subcaptionbox{Proposed\\ method: selective prior}{\includegraphics[width=0.24\columnwidth]{../images/tmh/RFA2/few_views/weighted_pca_all_methods_kk_0_01.png}}
\subcaptionbox{Weights map}{\includegraphics[width=0.275\columnwidth]{../images/tmh/RFA2/few_views/weights_colormap.png}}
\caption[Representative results-2]{\textbf{Goal: Observe details of new changes.} Reconstruction of slice 8 of Fig.~\ref{fig:RFA2_test_object-prior} from 120 views, using (b) FBP with SSIM $= 0.50$ (c) l1-ls with SSIM $=0.46$ and (d) unselective global prior, with SSIM $=0.51$ (\emph{notice dominance of the prior: a prominent residual shadow of the needle which was present in the object-prior, but not present in the test image}), and (e) selective prior with SSIM $=0.56$ (\emph{notice that the dominance of the prior is significantly controlled}). The region enclosed in green rectangle is our Region of Interest (RoI) as it contains both the new position of the needle and some background. The SSIM is computed in this RoI. (f) shows the computed weights map (defined later in the paper) used for reconstruction. Darker intensities indicate lower weights to prior as these are the regions of new changes.}
\label{fig:RFA2_few_views}
\end{figure}

@@

To begin with, when an object is scanned multiple times, a set of high quality reconstructions (i.e. reconstructions from a dense set of projection views) may be chosen as object-prior for the reconstruction of future scan volumes, which in turn, may be scanned using far fewer measurements. The eigenspace $E_{\text{high}}$ of the $L$ previously scanned objects $Q_1,Q_2,...,Q_L$ is pre-computed. Here, it is assumed that the test volume can be expressed as a sparse linear combination of the principal components (eigenvectors of the covariance matrix) obtained from a group of structurally similar volumes. Hence, the object-prior is represented by means of PCA. For the eigenspace to encompass a range of possible structures in the test slice, the object-prior must represent a wide structural range. Moreover, if these volumes are not aligned, then they must be first registered before computing the prior. The prior is built by computing the covariance matrix from the template set $\{Q_i\}_{i=1}^L$. \emph{The space spanned by the eigenvectors $\{\boldsymbol{V_k}\}_{k=1}^{L-1}$ (eigenspace) of the covariance matrix is the global prior and is assumed to contain any test slice that is similar, but not necessarily identical to the object-prior.}  We use all of the $L-1$ orthogonal eigenvectors as a basis to represent the unknown test volume. Let $\boldsymbol{\mu}$ denote the mean of the previously scanned objects, and $\boldsymbol{\alpha}$ the vector of eigen-coefficients of the test scan, of which $\alpha_k$ is the $k^{\textrm{th}}$ element. Then, once the eigenspace is pre-computed, the test is reconstructed by minimizing the following cost function:
\vspace{4mm}
 \begin{equation}
   \setlength{\belowdisplayskip}{0pt} \setlength{\belowdisplayshortskip}{0pt}
\setlength{\abovedisplayskip}{-2pt} \setlength{\abovedisplayshortskip}{-2pt}
J_1(\boldsymbol{\theta,\alpha}) = \lVert\boldsymbol{\mathcal{R} x}- \boldsymbol{y}\rVert_2^2  + \lambda_1\lVert\boldsymbol{\theta}\rVert_1+\lambda_2\lVert\boldsymbol{x} - (\boldsymbol{\mu} + \sum_{k}\boldsymbol{V_k}\alpha_k)\rVert_2^2.
\label{Eq:main_prior}
\end{equation}
Here, $\lambda_1$, $\lambda_2$ are tunable weights given to the sparsity and prior terms respectively. The unknowns $\boldsymbol{\theta}$ and $\boldsymbol{\alpha}$
are solved by alternately minimizing $J_{\boldsymbol{\alpha}}(\boldsymbol{\theta})$ using a fixed $\boldsymbol{\alpha}$, and $J_{\boldsymbol\theta}(\boldsymbol{\alpha})$ using the resultant $\boldsymbol{\theta}$, where 
\begin{align}
J_{\boldsymbol{\alpha}}(\boldsymbol{\theta}) &\triangleq \lVert\boldsymbol{\mathcal{R} x- y}\rVert_2^2  + \lambda_1\lVert\boldsymbol{\theta}\rVert_1+\lambda_2\lVert\boldsymbol{x} - (\boldsymbol{\mu + V\alpha})\rVert_2^2, \\
J_{\boldsymbol\theta}(\boldsymbol{\alpha}) &\triangleq \lVert\boldsymbol{\Upsilon\theta} - (\boldsymbol{\mu + V\alpha})\rVert_2^2.
\end{align}
Note that $\boldsymbol{\theta}$ is solved for using the basis pursuit CS solver~\cite{l1ls}. Solving for $\boldsymbol{\alpha}$ leads to the closed form update:
\begin{equation}
\boldsymbol{\boldsymbol{\alpha}} = \boldsymbol{V}^T(\boldsymbol{\Upsilon \theta} -\boldsymbol{\mu}).
\end{equation}
 Optimal values of $\lambda_1, \lambda_2$ must be empirically chosen \textit{a priori}, based on the reconstructions of one of the template volumes (see also Sec.~\ref{sec:discussion}). The cost function described in Eq.~\ref{Eq:main_prior} is biconvex and the convergence of this optimization is guaranteed by the monotone convergence theorem.
 %---------------------------------------------------------
 \vspace{2mm}

@@

%-----------------------------potato 2D weights
\begin{figure}[!h]
    \begin{subfigure}[b]{0.24\linewidth}
        \includegraphics[width=\textwidth]{../images/potato/2D/weightsIm_fbp30.png}
        \caption{FBP}
    \end{subfigure}
    \begin{subfigure}[b]{0.24\linewidth}
        \includegraphics[width=\textwidth]{../images/potato/2D/weightsIm_cs_dct30.png}
        \caption{(l1-ls)-DCT}
     \end{subfigure}
    \begin{subfigure}[b]{0.24\linewidth}
        \includegraphics[width=\textwidth]{../images/potato/2D/weightsIm_cs_wavelet30.png}
        \caption{(l1-ls)-Haar}
     \end{subfigure}
    \begin{subfigure}[b]{0.24\linewidth}
        \includegraphics[width=\textwidth]{../images/potato/2D/weightsIm_art30.png}
        \caption{ART}
     \end{subfigure}
    \begin{subfigure}[b]{0.24\linewidth}
        \includegraphics[width=\textwidth]{../images/potato/2D/weightsIm_sart30.png}
        \caption{SART}
     \end{subfigure}
    \begin{subfigure}[b]{0.24\linewidth}
        \includegraphics[width=\textwidth]{../images/potato/2D/weightsIm_sirt30.png}
        \caption{SIRT}
     \end{subfigure}
    \begin{subfigure}[b]{0.24\linewidth}
        \includegraphics[width=\textwidth]{../images/potato/2D/weightsIm_all_methods30.png}
        \caption{combined}
     \end{subfigure}
      \caption{Weights-maps (corresponding to the difference between pilot reconstruction of the image in the last sub-figure of Fig.~\ref{fig:object-prior_test_potato_A1} and its projection onto the eigenspace $V_{low}$) constructed using different reconstruction methods, as specified in Eq.~\ref{eq:weightsEq}. The weights-maps are different because the reconstruction artefacts of the new structures in test image will be different for every reconstruction method used, as seen in Fig.~\ref{fig:reconstructions_diff_methods}. }
\label{fig:weights_map_2Dpotato}
\end{figure}
%-----------------------------potato 2D weights
\begin{figure}[!h]
    \begin{subfigure}[b]{0.24\linewidth}
        \includegraphics[width=\textwidth]{../images/potato/2D/fbp.png}
        \caption{FBP}
    \end{subfigure}
    \begin{subfigure}[b]{0.24\linewidth}
        \includegraphics[width=\textwidth]{../images/potato/2D/cs_dct.png}
        \caption{(l1-ls)-DCT}
     \end{subfigure}
    \begin{subfigure}[b]{0.24\linewidth}
        \includegraphics[width=\textwidth]{../images/potato/2D/cs_wavelet.png}
        \caption{(l1-ls)-Haar}
     \end{subfigure}
    \begin{subfigure}[b]{0.24\linewidth}
        \includegraphics[width=\textwidth]{../images/potato/2D/art.png}
        \caption{ART}
     \end{subfigure}
    \begin{subfigure}[b]{0.24\linewidth}
        \includegraphics[width=\textwidth]{../images/potato/2D/sart.png}
        \caption{SART}
     \end{subfigure}
    \begin{subfigure}[b]{0.24\linewidth}
        \includegraphics[width=\textwidth]{../images/potato/2D/sirt.png}
        \caption{SIRT}
     \end{subfigure}
    \begin{subfigure}[b]{0.24\linewidth}
        \includegraphics[width=\textwidth]{../images/potato/2D/weighted_pca_all_methods30.png}
        \caption{combined}
     \end{subfigure}
      \caption{(a)-(f):Different reconstructions of~\ref{fig:potato_test}. The magnitude and sharpness of the artefacts is different for each method. (g) Selective-prior method that combines weights-map information from all other methods. The SSIM of these reconstructions is shown in Table.~\ref{table:potato_2D_ssim}.} 
\label{fig:reconstructions_diff_methods}
\end{figure}
